A hierarchical approach for modelling X-ray beamlines. Application to a coherent beamline.

### 2019-06-17 paper-hierarchical.tex
Updates from Overleaf

     %-------------------------------------------------------------------------
     % Information about journal to which submitted
     %-------------------------------------------------------------------------
     \journalcode{K}              % Indicate the journal to which submitted
                                  %   A - Acta Crystallographica Section A
                                  %   B - Acta Crystallographica Section B
                                  %   C - Acta Crystallographica Section C
                                  %   D - Acta Crystallographica Section D
                                  %   E - Acta Crystallographica Section E
                                  %   F - Acta Crystallographica Section F
                                  %   J - Journal of Applied Crystallography
                                  %   M - IUCrJ
                                  %   S - Journal of Synchrotron Radiation

@@

     % Use \shortauthor to indicate an abbreviated author list for use in
     % running heads (you will need to uncomment it).

@@

\begin{table}\label{tab:Undulators}
\centering
\caption{Parameters of the undulators available at the ID16A beamline: period ($\lambda_u$), length ($L$), number of identical undulators ($N$) and deflection parameter ($K$) values for the three main photon energies in use. The valuse are shown for the two ESRF magnetic lattices: the old ESRF-1 (labelled ESRF, with $E_e = 6.04$~GeV) and the new ESRF-EBS (labelled EBS, with $E_e = 6.00$~GeV). $n$ is the emission harmonic in use for each particular photon energy.}

\resizebox{\textwidth}{!}{%
\begin{tabular}{lcccccc}
name  & $\lambda_u$ [mm] & $L$ [m] & N & $K$ (11200~eV) & $K$ (17050~eV) & $K$ (33600~eV) \\ 
\input{tableUndulator.txt}
\end{tabular}%
}
\end{table}

@@

\begin{table}\label{tab:eBeam}
    \centering
    \caption{ESRF-1 and EBS main electron beam parameters. Electron beam parameters for the high-$\beta$ (labelled ESRF) and the EBS straight sections. Values taken at the symmetry point of the straight section, where the insertion devices are placed. Values for ESRF-1 are taken from \cite{ESRF2014}. For EBS, two (very close) design values are used, from Ref.~\cite{ESRF2014} labelled EBS-S28A, and those coming from further lattice refinement, labelled EBS-S28D. \todo{Remove ESRF-AT and EBS-S28A} }
        \resizebox{\textwidth}{!}{%
        \begin{tabular}{ccccc}
         Lattice & $\sigma_e$ horizontal {[}$\mu$m{]} &  $\sigma_e'$ horizontal {[}$\mu$rad{]} & $\sigma_e$ vertical {[}$\mu$m{]} & $\sigma_e'$ vertical {[}$\mu$rad{]} \\
         \input{tableElectron.txt}

%    \begin{tabular}{rcc}
%                                                & \textbf{ESRF (high-$\beta$)} & \textbf{EBS}     \\ 
%         $\sigma_e$ horizontal {[}$\mu$m{]}:    & 415.00                      & 30.20             \\
%         $\sigma_e'$ horizontal {[}$\mu$rad{]}: & 10.30                       & 4.40              \\
%         $\sigma_e$ vertical {[}$\mu$m{]}:      & 3.40                        & 3.60              \\
%         $\sigma_e'$ vertical {[}$\mu$rad{]}:   & 1.20                        & 1.40              \\ 
    \end{tabular}
    }
\end{table}

@@

\begin{figure}
    \centering
        \includegraphics[width=0.95\textwidth]{GRAPHICS/se_idealTE_idealOE_h.pdf}
        \includegraphics[width=0.95\textwidth]{GRAPHICS/ebs_slopeTE_slopeOE_h.pdf}
    \label{fig:SingleElectron}
    \caption{Intensity profile at the sample position produced by the single-electron emission propagated through the ID16A beamline. Top: the perfect focusing with ideal lenses (black) is compared with the focusing using elliptical mirrors (red). Bottom: the same as in the previous plot, but adding the effect of the mirrors slope errors in both cases.}
    %Simulations took less than one minute using a desktop computer Intel Core i7 with 8Gb of RAM memory.}
\end{figure}

@@

\begin{table}\label{tab:SRW_results}
\centering
\caption{Calculated beam dimensions (FWHM, in nm) at the focal position for different models of focusing elements using SRW.}
\resizebox{\textwidth}{!}{%
\begin{tabular}{ccccc}
Focusing elements                 & Plane      &  Single electron & EBS         & ESRF-1       \\ \\
ideal lenses/ elliptical mirrors  & Horizontal &  7.94/7.97       & 12.20/12.70 & 17.50/16.80 \\
the same, with errors             & Horizontal &  9.35/10.50       & 12.90/13.65 & 17.70/17.03 \\
ideal lenses/ elliptical mirrors  & Vertical   &  7.11/7.26       & 8.40/9.02   & 8.24/8.77    \\
the same, with errors             & Vertical   &  7.44/8.46       & 8.66/9.87   & 8.52/9.66    \\                                            
\end{tabular}
}
\end{table}

@@

\begin{figure}
    \centering
        \includegraphics[width=0.95\textwidth]{GRAPHICS/esrf_idealTE_idealOE_h.pdf}
        \includegraphics[width=0.95\textwidth]{GRAPHICS/esrf_slopeTE_slopeOE_h.pdf}
    \label{fig:FiniteEmittanceA}
    \caption{
    Intensity profile at the sample position produced by multi-electron method sampled with parameters of the ESRF-1 lattice, using ideal focusing elements (top), and including the effect of the mirrors slope errors (bottom).
    }
\end{figure}

@@

\begin{figure}
    \centering
        \includegraphics[width=0.95\textwidth]{GRAPHICS/ebs_idealTE_idealOE_h.pdf}
        \includegraphics[width=0.95\textwidth]{GRAPHICS/ebs_slopeTE_slopeOE_h.pdf}
    \label{fig:FiniteEmittanceB}
    \caption{
    Intensity profile at the sample position produced by multi-electron method sampled with parameters of the EBS lattice, using ideal focusing elements (top), and including the effect of the mirrors slope errors (bottom).
    }
\end{figure}

The results presented in this section are calculated using the Monte Carlo multi-electron method. It has been proposed by O. Chubar and it is fully implemented in SRW \cite{multiE}. The simulations presented here are built upon the ones described in the previous section (\ref{srw_se}). The different initial conditions were sampled from the 6D-electron phase space: its position on the cartesian space $(x_e, y_e, z_e)$; its direction in both vertical and horizontal planes  $(x'_e, y'_e)$ and its energy (as the electron energy spread is taken into account). Although simulations for highly coherent beamline converge rapidly (a few electrons are sufficient), this is in no way the case of the ID16A at ESRF-1: the VSS and the KB aperture reject a very high number of photons. To ensure good statistics when averaging the intensities of the different wavefronts from the sampled electrons, one has to go to a very high number of iterations. For the simulations presented here, 50000 wavefronts were used, amounting to a $\sim7$~h simulation time on a 56-core computer cluster. Convergence is faster for the EBS lattice as the electron-beam phase space is smaller.  Results are presented in Figures \ref{fig:FiniteEmittanceA} and \ref{fig:FiniteEmittanceB} and Table \ref{tab:SRW_results}.

@@

The combination of ray-tracing and wave optics methods used indicate that the focalized beam size due to  demagnification with mirrors (giving for EBS 10$\times$5~nm$^2$ analytically, 8.8$\times$4.5~nm$^2$ with ray-tracing) has to be corrected with diffraction effect (11.2$\times$8~nm$^2$ hybrid, inred{12.7$\times$9.2~nm$^2$ SRW}) and then add the slope error effect. The slope errors do not blow up the size and kit it in the specifications range (11.4$\times$8.5~nm$^2$ hybrid \inred{13.6$\times$9.9~nm$^2$ SRW}). It is important to treat the slope errors correctly. Slope errors can be included in the ray-tracing using the data in Appendix~\ref{appendix:metrology} and applying specular reflection on the optical surface with errors. This is the standard model of including slope errors in SHADOW, using measured profiles or synthetic ones. It works well for the case of incoherent beams, but in our case, because of the high coherence of the beam the ray-tracing method (specular reflection) overestimates the blowing of the focal spot conducting to wrong results. Therefore the slope errors of the KB system have been considered using a ``wave optics approach'', as explained in Ref.~\cite{hybrid}. Here, the mirror height errors $z(x)$ are projected to the plane perpendicular to the propagation ($z(x) \sin \theta$, with $\theta$ the grazing angle). They introduce a phase shift $\psi$ in the wavefront phase proportional to the optical path $\psi = k z(x) \sin \theta $. This is the algorithm used in the simulations in Sections~\ref{level1}, \ref{srw_se} and \ref{srw_me}.

>>>>

     %-------------------------------------------------------------------------
     % Information about journal to which submitted
     %-------------------------------------------------------------------------
     \journalcode{S}              % Indicate the journal to which submitted
                                  %   A - Acta Crystallographica Section A
                                  %   B - Acta Crystallographica Section B
                                  %   C - Acta Crystallographica Section C
                                  %   D - Acta Crystallographica Section D
                                  %   E - Acta Crystallographica Section E
                                  %   F - Acta Crystallographica Section F
                                  %   J - Journal of Applied Crystallography
                                  %   M - IUCrJ
                                  %   S - Journal of Synchrotron Radiation

@@

% emails:
%  srio@esrf.eu
%  celestre@esrf.eu
%  mark.glass@gmx.de
%  giogio.92@gmail.com
% juan.reyes-herrera@esrf.eu
% barrett@esrf.eu
% jdasilva@esrf.eu
% cloetens@esrf.eu
% xshi@anl.gov
% lrebuffi@anl.gov
%
%
%
     % Use \shortauthor to indicate an abbreviated author list for use in
     % running heads (you will need to uncomment it).

@@

\begin{table}\label{tab:Undulators}
\centering
\caption{Parameters of the undulators available at the ID16A beamline: period ($\lambda_u$), length ($L$), number of identical undulators ($N$) and deflection parameter ($K$) values for the three main photon energies in use. The valuse are shown for the two ESRF magnetic lattices: the old ESRF-1, with $E_e = 6.04$~GeV), and the new ESRF-EBS, with $E_e = 6.00$~GeV. $n$ is the emission harmonic in use for each particular photon energy.}

\resizebox{\textwidth}{!}{%
\begin{tabular}{lcccccc}
name  & $\lambda_u$ [mm] & $L$ [m] & N & $K$ (11200~eV) & $K$ (17050~eV) & $K$ (33600~eV) \\ \\
%\input{tableUndulator.txt}
 ESRF-1 U18.3  & 18.30  & 1.40  & 2 & 1.175(n=1) & 0.470(n=1) & 1.175(n=3)\\ 
 ESRF-1 U22.4  & 22.40  & 1.40  & 2 & 0.873(n=1) & 1.855(n=3) & 0.873(n=3)\\ \\
 EBS U18.3  & 18.30  & 1.40  & 2 & 1.156(n=1) & 0.437(n=1) & 1.156(n=3)\\ 
 EBS U22.4  & 22.40  & 1.40  & 2 & 0.852(n=1) & 1.836(n=3) & 0.852(n=3)\\ 
\end{tabular}%
}
\end{table}

@@

\begin{table}\label{tab:eBeam}
    \centering
    \caption{ESRF-1 and EBS main electron beam parameters. Electron beam parameters for the high-$\beta$ (labelled ESRF) and the EBS straight sections. Values taken at the symmetry point of the straight section, where the insertion devices are placed. Values are taken from \cite{ESRF2014}, updated for EBS with
    the lattice labelled S28D. 
    %For EBS, two (very close) design values are used, from Ref.~\cite{ESRF2014} labelled EBS-S28A, and those coming from further lattice refinement, labelled EBS-S28D. 
    %\todo{Remove ESRF-AT and EBS-S28A} 
    }
        \resizebox{\textwidth}{!}{%
        \begin{tabular}{ccccc}
         Lattice & $\sigma_e$ horizontal {[}$\mu$m{]} &  $\sigma_e'$ horizontal {[}$\mu$rad{]} & $\sigma_e$ vertical {[}$\mu$m{]} & $\sigma_e'$ vertical {[}$\mu$rad{]} \\ \\
         %\input{tableElectron.txt}
         ESRF-1  & 387.80  & 3.50  & 10.30 & 1.20 \\ 
         %ESRF-AT  & 414.97  & 3.43  & 10.31 & 1.16 \\ 
         %EBS-S28A  & 27.20  & 3.40  & 5.20 & 1.40 \\ 
         EBS     & 30.18  & 3.64  & 4.37 & 1.37 \\ 
    \end{tabular}
    }
\end{table}

@@

\begin{figure}
    \centering
        \includegraphics[width=0.95\textwidth]{GRAPHICS/filament_beam_ideal.png}
        \includegraphics[width=0.95\textwidth]{GRAPHICS/filament_beam_real.png}
    \label{fig:SingleElectron}
    \caption{Intensity profile at the sample position produced by the single-electron emission propagated through the ID16A beamline. Top: the perfect focusing with ideal lenses (green) is compared with the focusing using elliptical mirrors (blue). Bottom: the same as in the previous plot, but adding the effect of the mirrors slope errors in both cases.}
    %Simulations took less than one minute using a desktop computer Intel Core i7 with 8Gb of RAM memory.}
\end{figure}

@@

\begin{table}\label{tab:SRW_results}
\centering
\caption{Calculated beam dimensions (FWHM, in nm) at the focal position for different models of focusing elements using SRW.}
\resizebox{\textwidth}{!}{%
\begin{tabular}{ccccc}
Focusing elements                 & Plane      &  Single electron & EBS         & ESRF-1       \\ \\
ideal lenses/ elliptical mirrors  & Horizontal &  8.53/9.54       & 12.71/11.56 & 17.75/12.41 \\
the same, with errors             & Horizontal &  9.36/10.61      & 13.20/12.52 & 17.78/15.71 \\
ideal lenses/ elliptical mirrors  & Vertical   &  7.11/7.26       & 8.40/9.02   & 8.24/8.77    \\
the same, with errors             & Vertical   &  7.44/8.46       & 8.66/9.87   & 8.52/9.66    \\                                            
\end{tabular}
}
\end{table}

@@

\begin{figure}
    \centering
        \includegraphics[width=0.95\textwidth]{GRAPHICS/ESRF_ideal.png}
        \includegraphics[width=0.95\textwidth]{GRAPHICS/ESRF_real.png}
    \label{fig:FiniteEmittanceA}
    \caption{
    Intensity profile at the sample position produced by multi-electron method sampled with parameters of the ESRF-1 lattice, using ideal focusing elements (top), and including the effect of the mirrors slope errors (bottom).
    }
\end{figure}

@@

\begin{figure}
    \centering
        \includegraphics[width=0.95\textwidth]{GRAPHICS/EBS_ideal.png}
        \includegraphics[width=0.95\textwidth]{GRAPHICS/EBS_real.png}
    \label{fig:FiniteEmittanceB}
    \caption{
    Intensity profile at the sample position produced by multi-electron method sampled with parameters of the EBS lattice, using ideal focusing elements (top), and including the effect of the mirrors slope errors (bottom).
    }
\end{figure}

The results presented in this section are calculated using the Monte Carlo multi-electron method. It has been proposed by O. Chubar and it is fully implemented in SRW \cite{Chubar2011b}. The simulations presented here are built upon the ones described in the previous section (\ref{srw_se}). The different initial conditions were sampled from the 6D-electron phase space: its position on the cartesian space $(x_e, y_e, z_e)$; its direction in both vertical and horizontal planes  $(x'_e, y'_e)$ and its energy (as the electron energy spread is taken into account). Although simulations for highly coherent beamline converge rapidly (a few electrons are sufficient), this is in no way the case of the ID16A at ESRF-1: the VSS and the KB aperture reject a very high number of photons. To ensure good statistics when averaging the intensities of the different wavefronts from the sampled electrons, one has to go to a very high number of iterations. For the simulations presented here, 50000 wavefronts were used, amounting to a $\sim7$~h simulation time on a 56-core computer cluster. Convergence is faster for the EBS lattice as the electron-beam phase space is smaller.  Results are presented in Figures \ref{fig:FiniteEmittanceA} and \ref{fig:FiniteEmittanceB} and Table \ref{tab:SRW_results}.

@@

The combination of ray-tracing and wave optics methods used indicate that the focalized beam size due to  demagnification with mirrors (giving for EBS 10$\times$5~nm$^2$ analytically, 8.8$\times$4.5~nm$^2$ with ray-tracing) has to be corrected with diffraction effect (11.2$\times$8~nm$^2$ hybrid, 12.7$\times$9.2~nm$^2$ SRW) and then add the slope error effect. The slope errors do not blow up the size and kit it in the specifications range (11.4$\times$8.5~nm$^2$ hybrid \inred{13.6$\times$9.9~nm$^2$ SRW}). It is important to treat the slope errors correctly. Slope errors can be included in the ray-tracing using the data in Appendix~\ref{appendix:metrology} and applying specular reflection on the optical surface with errors. This is the standard model of including slope errors in SHADOW, using measured profiles or synthetic ones. It works well for the case of incoherent beams, but in our case, because of the high coherence of the beam the ray-tracing method (specular reflection) overestimates the blowing of the focal spot conducting to wrong results. Therefore the slope errors of the KB system have been considered using a ``wave optics approach'', as explained in Ref.~\cite{hybrid}. Here, the mirror height errors $z(x)$ are projected to the plane perpendicular to the propagation ($z(x) \sin \theta$, with $\theta$ the grazing angle). They introduce a phase shift $\psi$ in the wavefront phase proportional to the optical path $\psi = k z(x) \sin \theta $. This is the algorithm used in the simulations in Sections~\ref{level1}, \ref{srw_se} and \ref{srw_me}.
